---
title: "Peer Assessment1"
author: "Muthiah Sriram"
date: "Sunday, May 17, 2015"
output: html_document
---

```{r setoptions ,echo =TRUE,warning=FALSE}
library(knitr)
opts_chunk$set(warning = FALSE)


```

```{r question1,echo =T,cache=FALSE}
setwd("C:\\Studies\\Coursera_Ds_reproducible research\\repdata_data_activity")
data = read.csv("activity.csv")
#removing NAs
good = complete.cases(data)
data1 = data[good,]
#Date
data1$date = as.Date(data1$date ,format = "%Y-%m-%d" ) 
#histogram
s = split(data1,data1$date)
steps_summary = as.data.frame(sapply(s,function(x) sum(x[,c("steps")])))
steps_summary = cbind(steps_summary , rownames(steps_summary))
colnames(steps_summary) = c("steps","date")
mean = mean(steps_summary$steps)
median = median(steps_summary$steps)
hist(steps_summary$steps,breaks = 20,col = "blue",xlab = "steps/day",main = "histogram of steps/day")


```

-The mean of totalNumberOfSteps/day is _`r mean`_

-The median of totalNumberOfSteps/day is _`r median`_



```{r timeseries plot,echo =TRUE}
s2 = split(data1,data1$interval)
interval_summary =as.data.frame(sapply(s2,function(x){mean(x[,c("steps")])}))
interval_summary = cbind(interval_summary , rownames(interval_summary))
colnames(interval_summary) = c("steps","interval")

interval_summary$interval = factor(interval_summary$interval, ordered=TRUE ,levels = as.character(sort(as.numeric(levels(interval_summary$interval)))))
maxInterval = interval_summary$interval[which.max(interval_summary$steps)]
plot(interval_summary$interval,interval_summary$steps,main = "Timeseries plot ",xlab = "interval",ylab ="steps")

lines(interval_summary$interval,interval_summary$steps,type="l",col = "Blue")
abline(v =  interval_summary$interval[which.max(interval_summary$steps)],col ="red")
text(y = max(interval_summary$steps)-10 , x=interval_summary$interval[which.max(interval_summary$steps)] ,"interval with max steps",col = "green" )

```



-From the graph we can infer that the interval with the highest number of steps is _`r maxInterval`_
 

```{r imputing code}

#using the mean value in that interval
data12 = merge(data,interval_summary , by = "interval")
data12$date = as.Date(data12$date ,format = "%Y-%m-%d" ) 
data12 = data12[order(data12$date,data12$interval),]
data12[is.na(data12$steps.x),]$steps.x = data12[is.na(data12$steps.x),]$steps.y
data12 = data12[,c(1,2,3)]
colnames(data12) = c("interval","steps","date")

```

```{r , fig.height=8}
s12 = split(data12,data12$date)
steps_summary12 = as.data.frame(sapply(s12,function(x) sum(x[,c("steps")])))
steps_summary12 = cbind(steps_summary12 , rownames(steps_summary12))
colnames(steps_summary12) = c("steps","date")
mean1=mean(steps_summary12$steps)
median1=median(steps_summary12$steps)
par(mfrow=c(2,1))
hist(steps_summary12$steps,breaks = 20,col = "red",xlab = "steps/day",main = "After Imputing")
hist(steps_summary$steps,breaks = 20,col = "blue",xlab = "steps/day",main = "Before Imputing")


```



-The mean of totalNumberOfSteps/day after imputing is _`r mean1`_

-The median of totalNumberOfSteps/day after imputing is _`r median1`_

-Since we used the mean of the steps per interval for filling NA values,there is not much chnage in the mean and median.Only the frequency of values have changed in the histogram(as expected)





__Imputed Data__

```{r addingNewColumn}

data12$day = ifelse(weekdays(data12$date)=="Saturday"|weekdays(data12$date)=="Sunday" ,"Weekend","Weekday")
#  library(xtable)
#  table = xtable(head(data12,5))
print(head(data12,6))

```


```{r graphs,fig.height=8}
s3 = split(data12[data12$day == "Weekend",],data12$interval)
s4 = split(data12[data12$day == "Weekday",],data12$interval)

graph = function(x,words ="weekend",col){
interval_summary12 =as.data.frame(sapply(x,function(x){mean(x[,c("steps")])}))
interval_summary12 = cbind(interval_summary12 , rownames(interval_summary12))
colnames(interval_summary12) = c("steps","interval")

interval_summary12$interval = factor(interval_summary12$interval, ordered=TRUE ,levels = as.character(sort(as.numeric(levels(interval_summary12$interval)))))
plot(interval_summary12$interval,interval_summary12$steps,main = words)
lines(interval_summary12$interval,interval_summary12$steps,type="l",col = col,xlab = "interval","steps")
}

par(mfrow=c(2,1))
graph(s3,"Weekend","Blue")
graph(s4,"Weekday","Red")


```
